\documentclass[main.tex]{subfiles}

\begin{document}
I must first thank my advisor Dan Licata. Your guidance and understanding, both during your class and throughout my research under you, have been invaluable. Throughout my constant misunderstandings in the incredibly slow process that is learning, I never once felt judged but rather encouraged. Your clarity of thought, command of logic, and graceful attitude are deeply inspiring.  I am especially grateful for your patience during many meetings, even when my thoughts were scattered or incomplete. Talking with you often revealed how much more I had to learn, and those moments were pivotal for my growth. Thank you for having trust in me and allowing me autonomy in choosing my topic, I truly cannot overstate how working with you has changed the direction of my life for the better. 

I also extend my gratitude to Robert Rose. I appreciate all the knowledge sharing sessions we had, and you made me truly understand the level of work and effort required to understand things deeply. Watching you solve problems has allowed me to see that things do not always have to be difficult or scattered. Your diligence and astuteness left a lasting impression on me, and I aspire to embody these qualities in my own work one day.

Trey, I owe you thanks for being a great friend and a constant presence in the CS lab, where I spent countless hours. I’m incredibly grateful for our early work sessions, even when we were slightly lost on something as simple as a derivation tree. Reflecting on how far we’ve come brings me joy. Our conversations—ranging from AI alignment, random philosophy, and the news of the day to type theory and our respective projects—kept me grounded and motivated throughout this journey.

Finally, thank you, Selena, for being my rock. Your unwavering support, not just during this thesis but throughout college, has been a constant source of strength. You have been a true friend and an anchor for me. I genuinely could not have done this without you by my side. Thank you. A million thank yous.

\end{document}